### 2014-12-01 sharp-sigma-plus.tex
f-1 v.1

\begin{theo}
$\#\Sigma_1$ is closed under substraction of 1 element.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{z}) \} \vert
$$
our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
$$
F'(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
$$
where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$. Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{P}) \} \vert
$$
similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples. First we treat sub-formulas without $\vec{P}$ as literals, and take a DNF formula that's equivalent to $\varphi$. Then
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi_k(\vec{x},\vec{P})
 \} \vert
$$
Note that
$$
\varphi_i(\vec{x},\vec{P}) = \vec{Q}(\vec{x}) \wedge \neg \vec{R}(\vec{x}) \wedge \varphi'_i(\vec{x})
$$
where $\vec{Q},\vec{R}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified. We can assume that variables in $\vec{x}$ appear in at most one predicate since we can replace common variables with new ones and add appropriate equals in $\varphi'_i$. Hence, if there is at least one assignment to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal. Thus, if
$$
\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
$$
then 
$$
\exists \vec{x} \left( \varphi_i(\vec{x},\vec{P}) \wedge \psi_i(\vec{x}) \right)
$$
will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$. If $\varphi'_i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$. If this sub-formula is also unsatisfiable, we move to the next one, and so on. Therefore:
\begin{multline*}
F_{-1}(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
\exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
\exists \vec{x} \left( \varphi_2(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} \neg \varphi'_1(\vec{x'})) \rightarrow \psi_2(\vec{x}) \right)\right) \vee \\
... \vee
\exists \vec{x} \left( \varphi_k(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} (\neg \varphi'_1(\vec{x'}) \wedge \neg \varphi'_2(\vec{x'}) \wedge ... \wedge \neg \varphi'_{k-1}(\vec{x'})) ) \rightarrow \psi_k(\vec{x}) \right)\right)
 \} \vert
\end{multline}
eliminates exactly one predicate of $F(\A)$.
\end{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x}, \vec{P},\vec{z}) \} \vert
$$
Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it. We mix both previous strategies. For each DNF disjunct $\varphi_i(\vec{x}, \vec{z})$ we add:
$$
\psi_i(\vec{x}, \vec{z}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x)) \vee \exists z'(\varphi_i(\vec{x}, \vec{z}) \wedge \varphi_{\lneq}(z',z))
$$
and proceed analogously.
\end{proof}

>>>>

\begin{theo}
$\#\Sigma_1$ is closed under substraction of 1 element.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{z}) \} \vert
$$
our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
$$
F'(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
$$
where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$. Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{P}) \} \vert
$$
similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples. First we treat sub-formulas without $\vec{P}$ as literals, and take a DNF formula that's equivalent to $\varphi$. Then
$$
F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi_k(\vec{x},\vec{P})
 \} \vert
$$
We can also suppose that all of this formulas are satisfied by a single total order for $\vec{x}$. Therefore
$$
\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})
$$
where $P_i$ and $\overline{P}$ are subsequences of $\vec{P}$, $\vec{x_i}$ and $\vec{y}$ are subsequences of $\vec{x}$, $\theta$ defines a total order for $\vec{x}$, and $\varphi_{\vec{S}}$ involves only closed predicates in $\vec{S}$. We now define
\begin{multline*}
\alpha_{\min}^i(\vec{x_1}',...,\vec{x_n}') = \\ \exists\vec{y}\left( \alpha(\vec{x_1}',...,\vec{x_n}',\vec{y})\wedge \forall\vec{x_1}''...\vec{x_n}''\vec{y}'(\alpha(\vec{x_1}'',...,\vec{x_n}'',\vec{y}')\rightarrow (\vec{x_1}',...,\vec{x_n}') \leq (\vec{x_1}'',...,\vec{x_n}'')\right)
\end{multline}
Where $\alpha(\vec{x}) = \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})$. Our new formula would be
\begin{multline*}
\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})\wedge\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
\left(\bigvee_{P\in\overline{P}}\bigvee_{P_i = P}\neg P(\vec{x_i}') \vee \bigvee_{P\in\overline{P}} \exists \vec{x}''\left(P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\right)\right)
\end{multline}
and $\vec{x_1}',...,\vec{x_n}'$ are added to the existential quantifier. This will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.

If $\alpha^i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$. If this sub-formula is also unsatisfiable, we move to the next one, and so on. Therefore:

\end{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x}, \vec{P},\vec{z}) \} \vert
$$
Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it. We mix both previous strategies. For each DNF disjunct $\varphi_i(\vec{x}, \vec{z})$ we add:
$$
\psi_i(\vec{x}, \vec{z}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x)) \vee \exists z'(\varphi_i(\vec{x}, \vec{z}) \wedge \varphi_{\lneq}(z',z))
$$
and proceed analogously.
\end{proof}
